---
title: "Take Home Exercise 2 Part 3"
author: "Xu Haiyang"
date: "23 September, 2024" 
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---
```{r}
pacman::p_load(sf, dplyr, readr, ggplot2, sfdep, tmap, tidyverse)
```

```{r}
Drug <- read_csv("data/thai_drug_offenses_2017_2022.csv")
```

```{r}
thai_geo = st_read(dsn = "data/tha_adm_rtsd_itos_20210121", layer = "tha_admbnda_adm1_rtsd_20220121")
```

```{r}
thai_geo_projected <- st_transform(thai_geo, crs = 32647)
```

```{r}
thai_geo_projected <- thai_geo_projected %>%
  mutate(
    Shape_Area = st_area(geometry),
    Shape_leng = st_length(st_cast(geometry, "MULTILINESTRING"))
  )
```

```{r}
Drug <- read_csv("data/thai_drug_offenses_2017_2022.csv") %>%
  mutate(
    province_en = recode(province_en, 
                         "buogkan" = "Bueng Kan",
                         "Loburi" = "Lop Buri")
  )
```

```{r}
merged_data <- left_join(
  thai_geo_projected,
  Drug,
  by = c("ADM1_EN" = "province_en")
) %>%
  select(
    ADM1_EN,
    ADM1_PCODE,
    Shape_leng,
    Shape_Area,
    fiscal_year,
    types_of_drug_offenses,
    no_cases,
    geometry
  )

rm(Drug, thai_geo, thai_geo_projected)
```

```{r}
drug_use_cases_data <- merged_data %>%
  filter(fiscal_year == "2021", types_of_drug_offenses == "drug_use_cases")
```

```{r}
rm(merged_data)
```

```{r}
neighbors_drug_use <- st_contiguity(drug_use_cases_data, queen = TRUE)
summary(neighbors_drug_use)
```

```{r}
centroids_drug_use <- st_centroid(drug_use_cases_data)
  
neighbors_drug_use <- st_knn(centroids_drug_use, k = 5)
```

```{r}
weights_drug_use <- st_weights(neighbors_drug_use, style = "W")
```

```{r}
lisa_drug_use <- drug_use_cases_data %>% 
  mutate(local_moran = local_moran(
    no_cases,
    neighbors_drug_use,
    weights_drug_use, 
    nsim = 99),
    .before = 1) %>%
  unnest(local_moran)
lisa_drug_use
```

# **9. Emerging Hot Spots Analysis (EHSA)**

## **9.1 Local Getis-Ord** Gi\* for **Hot Spot and Cold Spot Area Analysis**

I tried to analyze the **spatial clustering** of drug use cases using the **Local Getis-Ord Gi\*** statistic. This analysis identifies specific areas that are statistically significant **hotspots** or **coldspots**, representing areas of unusually high or low concentrations of the phenomena.

### **9.1.1 Hotspot and Coldspot Analysis for Drug Use Cases**

I first calculated the spatial weights using inverse distance weighting (IDW) to account for proximity between areas. I then applied the **local Getis-Ord Gi\*** permutation test to detect significant hot and cold spots for drug use cases.

```{r}
wm_idw_drug_use <- lisa_drug_use %>%
  mutate(nb = include_self(st_contiguity(geometry)),
         wt = st_inverse_distance(nb, geometry, scale = 1, alpha = 1),
         .before = 1)
```

After calculating the spatial weights, I computed the **local Gi\*** for drug use cases:

```{r}
HCSA_drug_use <- wm_idw_drug_use %>% 
  mutate(local_Gi_star = local_gstar_perm(no_cases, neighbors_drug_use, weights_drug_use, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi_star, names_sep = "_") %>%
  mutate(label = paste(ADM1_EN))
```

Next, I visualized the resulting hot and cold spots for drug use cases. The color palette ranges from **coldspots** (blue) to **hotspots** (red), providing a clear view of spatial patterns.

```{r}
tmap_mode("plot")  
tm_shape(HCSA_drug_use) +
  tm_fill("local_Gi_star_gi_star",  # Use the actual name of the 'gi_star' column
          palette = c("#57bfc0", "#7977f3", "#f8d673", "#f8b675", "#f67774"),
          title = "Gi*",
          midpoint = 0) +
  tm_borders(col = "black", alpha = 0.6) +
  tm_layout(main.title = "Hotspots & Coldspots of Drug Use Cases",
            main.title.position = "center",
            main.title.size = 1.7,
            main.title.fontface = "bold",
            legend.title.size = 1.8,
            legend.text.size = 1.3,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", text.size = 1.5, size = 3, position = c("RIGHT", "TOP")) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), text.size = 1.2) +
  tm_grid(labels.size = 1, alpha = 0.2)
```

I filtered the dataset to focus on the statistically significant hot and cold spots (p \< 0.05).

```{r}
HCSA_sig_drug_use <- HCSA_drug_use %>% filter(p_ii_sim < 0.05)
```

The following map provides an interactive view of the significant hot and cold spots:

```{r}
tmap_mode("view")
tm_shape(HCSA_drug_use) +
  tm_polygons(id = "label") +
tm_shape(HCSA_sig_drug_use) +
  tm_fill("local_Gi_star_gi_star",
          palette = c("#57bfc0", "#7977f3", "#f8d673", "#f8b675", "#f67774"),
          title = "Gi*",
          midpoint = 0,
          id = "label") +
  tm_borders(col = "black", alpha = 0.6)
```

### **9.1.2 Identifying the Most Significant Hotspots and Coldspots for Drug Use Cases**

In this step, I identified the **three most significant hotspots** and **three most significant coldspots** for drug use cases based on their Gi\* values. These regions are where clustering of drug use cases is at its highest (hotspots) or lowest (coldspots), providing critical insights into geographical patterns of drug-related offenses.

First, I filtered for the most significant areas using Gi\* values:

```{r}
set.seed(123)
three_hotspots_drug_use <- (head((HCSA_sig_drug_use[HCSA_sig_drug_use$local_Gi_star_gi_star > -2, ]), 3))$label

three_coldspots_drug_use <- (head((HCSA_sig_drug_use[HCSA_sig_drug_use$local_Gi_star_gi_star < -2, ]), 3))$label
```

The **two most significant hotspots** are:

```{r}
three_hotspots_drug_use
```

The **two most significant coldspots** are:

```{r}
three_coldspots_drug_use
```

Next, I filtered the original dataset to highlight only these key areas:

```{r}
HCSA_three_hotspots_drug_use <- HCSA_sig_drug_use %>% filter(label %in% three_hotspots_drug_use)
HCSA_three_coldspots_drug_use <- HCSA_sig_drug_use %>% filter(label %in% three_coldspots_drug_use)
```

#### **9.1.2.1 Visualizing Hotspots**

I created a map to visualize the two most significant hotspots for drug use cases, where the areas with the strongest clustering of high values are highlighted:

```{r}
tmap_mode("plot")
tm_shape(HCSA_drug_use) +
  tm_polygons() +
tm_shape(HCSA_three_hotspots_drug_use) +
  tm_fill("local_Gi_star_gi_star", palette = "#f67774") +
  tm_borders(col = "black", alpha = 0.6) +
  tm_text("label", auto.placement = TRUE) +
  tm_layout(main.title = "Two Most Significant Hotspots of Drug Use Cases",
            main.title.position = "center",
            main.title.size = 1.7,
            main.title.fontface = "bold",
            legend.show = FALSE,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", text.size = 1.5, size = 3, position = c("RIGHT", "TOP")) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), text.size = 1.2) +
  tm_grid(labels.size = 1, alpha = 0.2)
```

#### **9.1.2.2 Visualizing Coldspots**

Similarly, I visualized the two most significant coldspots, which highlight areas with the lowest clustering of drug use cases:

```{r}
tm_shape(HCSA_drug_use) +
  tm_polygons() +
tm_shape(HCSA_three_coldspots_drug_use) +
  tm_fill("local_Gi_star_gi_star", palette = "#57bfc0") +
  tm_borders(col = "black", alpha = 0.6) +
  tm_text("label", auto.placement = TRUE) +
  tm_layout(main.title = "Two Most Significant Coldspots of Drug Use Cases",
            main.title.position = "center",
            main.title.size = 1.7,
            main.title.fontface = "bold",
            legend.show = FALSE,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", text.size = 1.5, size = 3, position = c("RIGHT", "TOP")) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), text.size = 1.2) +
  tm_grid(labels.size = 1, alpha = 0.2)
```

# **10. Analysis Summary and Conclusions**

This study has employed a comprehensive suite of spatial autocorrelation and clustering techniques to investigate the geographic patterns of drug use cases across provinces in Thailand. These methods provide critical insights into the spatial dynamics of drug-related activities, revealing significant clusters, outliers, and emerging trends.

## **Global Spatial Autocorrelation Analysis**

The global measures of spatial autocorrelation, including **Moran’s I** and **Geary’s C**, indicated that both drug use cases and suspects in drug use cases are not randomly distributed across Thai provinces. The positive Moran's I values suggest a significant level of spatial clustering, where provinces with high or low levels of drug use cases or suspects tend to be surrounded by provinces with similar characteristics.

-   **Moran's I** revealed significant clustering of similar values, with provinces exhibiting high drug-related cases surrounded by other high-value provinces and similarly for low-value provinces. This was confirmed through both the observed statistic and Monte Carlo permutation tests.

-   **Geary’s C**, being more sensitive to local variations, complemented Moran’s I by identifying finer details of spatial autocorrelation, such as the presence of spatial outliers or smaller clusters that may have been overlooked in a global analysis.

## **Local Spatial Autocorrelation (LISA) and Hotspot Analysis**

The **Local Moran’s I (LISA) analysis** provided a more granular view, revealing specific provinces where clustering of drug use cases is statistically significant. Through LISA, we classified provinces into high-high clusters (hotspots), low-low clusters (coldspots), and identified outliers where low values are surrounded by high values (low-high) and vice versa (high-low).

-   **Hotspots**: Provinces such as **Amnat Charoen** and **Si Sa Ket** were identified as high-high clusters for both drug use cases, indicating regions where the incidence of drug-related activities is highly concentrated. These areas could be prioritized for targeted interventions or resources.

-   **Coldspots**: Provinces like **Kamphaeng Phet** and **Nakhon Sawan** formed low-low clusters, indicating regions where drug-related activities are significantly less concentrated. This might suggest successful prevention strategies or simply less drug-related activity in these regions.

-   **Outliers**: Notable outliers such as **Krabi**, which exhibited a low-high pattern, indicated that this province has lower-than-expected drug use cases but is surrounded by higher-value provinces. These outliers can be critical for identifying areas that require further investigation to understand local factors influencing drug activity.

## **Local Getis-Ord Gi\* and Emerging Hot Spots Analysis (EHSA)**

The **Local Getis-Ord Gi\*** statistic provided additional insights into clustering by identifying specific provinces where high or low values are spatially concentrated. This analysis complemented LISA by offering a more robust identification of statistically significant hotspots and coldspots for drug-related activities.

-   **Hotspots for Drug Use Cases**: Provinces like **Ubon Ratchathani** and **Si Sa Ket** were identified as the two most significant hotspots for drug use cases. These regions demonstrate a strong spatial concentration of high values, making them critical focal points for policy interventions.

-   **Coldspots for Drug Use Cases**: Conversely, provinces like **Kamphaeng Phet** and **Nakhon Sawan** were highlighted as the two most significant coldspots, showcasing areas with significantly low levels of drug use cases.

**Key Findings and Policy Implications**

1.  **Concentration of Drug-Related Activity in Specific Regions**: The consistent identification of provinces such as **Amnat Charoen**, **Si Sa Ket**, and **Ubon Ratchathani** as hotspots suggests that these areas are major centers of drug-related activities. These provinces should be prioritized for targeted interventions such as increased law enforcement presence, drug prevention programs, and community outreach initiatives.

2.  **Need for Tailored Strategies in Coldspot Regions**: The low-low clusters and coldspots, such as **Kamphaeng Phet** and **Nakhon Sawan**, indicate regions with lower levels of drug-related activity. While these areas may not require immediate intervention, understanding the factors that contribute to their low rates could inform successful drug prevention strategies in other parts of the country.

3.  **Emerging Trends in Outlier Regions**: The presence of spatial outliers, particularly provinces like **Krabi** (low-high pattern), underscores the importance of localized factors that may influence drug use patterns. Understanding these outliers can help develop tailored strategies that address unique local conditions, such as regional socioeconomic factors, infrastructure, or law enforcement practices.

## **Conclusion**

The application of spatial autocorrelation techniques, including global Moran’s I, Geary’s C, Local Moran’s I (LISA), and Local Getis-Ord Gi\*, has allowed for a detailed understanding of the spatial distribution of drug-related activities in Thailand. This study has demonstrated that drug use cases and suspects are not randomly distributed, but instead, form distinct clusters of high and low activity across provinces.

The identification of **hotspots** and **coldspots** provides valuable insights for policymakers, allowing for the allocation of resources and the implementation of targeted interventions in regions where drug-related activity is most concentrated. Furthermore, the detection of **spatial outliers** highlights the need for localized strategies that consider unique regional factors.

By leveraging these spatial analysis techniques, authorities can develop data-driven strategies that effectively address the drug problem in Thailand, focusing on prevention, enforcement, and rehabilitation in areas where they are most needed.
