---
title: "Take Home Exercise 2 Part 2"
author: "Xu Haiyang"
date: "23 September, 2024" 
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

```{r}
pacman::p_load(sf, dplyr, readr, ggplot2, sfdep, tmap, tidyverse)
```

```{r}
Drug <- read_csv("data/thai_drug_offenses_2017_2022.csv")
```

```{r}
thai_geo = st_read(dsn = "data/tha_adm_rtsd_itos_20210121", layer = "tha_admbnda_adm1_rtsd_20220121")
```

```{r}
thai_geo_projected <- st_transform(thai_geo, crs = 32647)
```

```{r}
thai_geo_projected <- thai_geo_projected %>%
  mutate(
    Shape_Area = st_area(geometry),
    Shape_leng = st_length(st_cast(geometry, "MULTILINESTRING"))
  )
```

```{r}
Drug <- read_csv("data/thai_drug_offenses_2017_2022.csv") %>%
  mutate(
    province_en = recode(province_en, 
                         "buogkan" = "Bueng Kan",
                         "Loburi" = "Lop Buri")
  )
```

```{r}
merged_data <- left_join(
  thai_geo_projected,
  Drug,
  by = c("ADM1_EN" = "province_en")
) %>%
  select(
    ADM1_EN,
    ADM1_PCODE,
    Shape_leng,
    Shape_Area,
    fiscal_year,
    types_of_drug_offenses,
    no_cases,
    geometry
  )

rm(Drug, thai_geo, thai_geo_projected)
```

```{r}
drug_use_cases_data <- merged_data %>%
  filter(fiscal_year == "2021", types_of_drug_offenses == "drug_use_cases")
```

```{r}
rm(merged_data)
```

```{r}
neighbors_drug_use <- st_contiguity(drug_use_cases_data, queen = TRUE)
summary(neighbors_drug_use)
```

```{r}
centroids_drug_use <- st_centroid(drug_use_cases_data)
  
neighbors_drug_use <- st_knn(centroids_drug_use, k = 5)
```

```{r}
weights_drug_use <- st_weights(neighbors_drug_use, style = "W")
```

# **8. Local Moran’s I Test**

I tried to compute the **Local Moran's I** statistic for drug use cases, which allows for the identification of local clusters and spatial outliers. While the **Global Moran’s I** test examines overall spatial autocorrelation, **Local Moran’s I** focuses on the spatial dependence within smaller areas, helping to identify **hotspots**, **coldspots**, and potential outliers in specific provinces.

## **8.1 Calculating Local Moran’s I for Drug Use Cases**

I calculated Local Moran’s I for drug use cases in 2021. By using 99 permutations, I can assess the significance of local spatial autocorrelation and detect areas where drug use cases cluster more than expected under randomness.

```{r}
lisa_drug_use <- drug_use_cases_data %>% 
  mutate(local_moran = local_moran(
    no_cases,
    neighbors_drug_use,
    weights_drug_use, 
    nsim = 99),
    .before = 1) %>%
  unnest(local_moran)
lisa_drug_use
```

## **8.2 Visualizing Local Moran’s I Results**

I tried to visualize the **Local Moran’s I values** for drug use cases. This visualization helps to understand the spatial clustering and identify potential hotspots or coldspots.

### **8.2.1 Visualizing Local Moran’s I for Drug Use Cases**

The following map shows the Local Moran’s I values for drug use cases, highlighting areas of high or low spatial autocorrelation. Provinces with significant spatial clusters or outliers are emphasized using a color palette ranging from blue (low) to red (high).

```{r}
tmap_mode("plot")

tm_shape(lisa_drug_use) +
  tm_fill(
    "ii", 
    palette = c("#b7dce9", "#e1ecbb", "#f5f3a6", "#f8d887", "#ec9a64", "#d21b1c"), 
    title = "Local Moran's I",
    midpoint = NA,
    legend.hist = TRUE,
    legend.is.portrait = TRUE,
    legend.hist.z = 0.1
  ) +
  tm_borders(col = "black", alpha = 0.6) +
  tm_layout(
    main.title = "Spatial Autocorrelation of Drug Use Cases (Local Moran's I)",
    main.title.position = "center",
    main.title.size = 1.7,
    main.title.fontface = "bold",
    legend.title.size = 1.8,
    legend.text.size = 1.3,
    frame = TRUE
  ) +
  tm_borders(alpha = 0.5) +
  tm_compass(
    type = "8star", text.size = 1.5, size = 3, position = c("RIGHT", "TOP")
  ) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), text.size = 1.2) +
  tm_grid(labels.size = 1, alpha = 0.2)
```

In the map showing **Local Moran's I** results for drug use cases across Thailand, regions are color-coded to represent spatial clusters of similar values. The **color gradient** ranges from **blue** (indicating regions with lower values of Local Moran's I) to **red** (indicating regions with higher positive spatial autocorrelation). This implies that:

-   **Red areas** (in the northeast and parts of the southeast) represent **hotspots**, where regions have high values of drug use and are surrounded by neighbors with similarly high values.

-   **Yellow and light orange areas** represent regions with lower clustering or random patterns of drug use, indicating a more neutral or less significant spatial autocorrelation.

-   **Blue areas** in the southern part of Thailand indicate **coldspots**, where regions with low drug use are surrounded by neighbors with similarly low values.

The **north-south** divide in spatial autocorrelation is apparent, with more clustering of high drug use in the northern and northeastern regions and lower clustering in the southern regions. This map highlights where spatial patterns of drug use are concentrated, which could inform targeted interventions or policies.

## **8.3 Visualizing the Statistical Significance of Local Moran’s I p-values**

In addition to visualizing the Local Moran’s I values, I also tried to map the **p-values** for the drug use cases. This helps identify areas where the spatial autocorrelation is statistically significant, providing more context on how likely it is that observed clustering occurred by chance.

### **8.3.1 p-value Visualization for Drug Use Cases**

The map below visualizes the statistical significance of Local Moran’s I for drug use cases, where lower p-values indicate more significant clustering.

```{r}
tm_shape(lisa_drug_use) +
  tm_fill(
    "p_ii_sim", 
    palette = c("#b7dce9", "#c9e3d2", "#f5f3a6", "#ec9a64", "#d21b1c"), 
    title = "p-value",
    midpoint = NA,
    legend.hist = TRUE, 
    legend.is.portrait = TRUE,
    legend.hist.z = 0.1
  ) +
  tm_borders(col = "black", alpha = 0.6) +
  tm_layout(
    main.title = "Statistical Significance of Spatial Autocorrelation\nof Drug Use Cases",
    main.title.position = "center",
    main.title.size = 1.7,
    main.title.fontface = "bold",
    legend.title.size = 1.8,
    legend.text.size = 1.3,
    frame = TRUE
  ) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", text.size = 1.5, size = 3, position = c("RIGHT", "TOP")) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), text.size = 1.2) +
  tm_grid(labels.size = 1, alpha = 0.2)
```

In this map, representing the **p-values** for the **Local Moran's I** test on **drug use cases**, the regions are color-coded based on the statistical significance of their spatial autocorrelation. The **p-value** indicates the likelihood that the observed spatial pattern is due to random chance, with lower values signifying greater statistical significance:

-   **Red areas** indicate regions where the p-value is very low (close to zero), suggesting that the spatial clustering of cases is statistically significant and not due to chance. These regions are likely **hotspots** with a high number of cases in close proximity to other high-suspect areas.

-   **Blue areas** represent higher p-values (closer to 1), indicating regions where the spatial pattern is less statistically significant or random, meaning no strong spatial clustering of cases.

-   **Yellow and light orange regions** fall somewhere in between, showing moderate significance.

This map helps to identify areas where the spatial clustering of drug use cases is most statistically significant, especially in the northern and central provinces. These regions may require more attention for addressing drug-related issues, as the spatial patterns are unlikely to be random.

## **8.4 Visualising Statistically Significant Local Spatial Autocorrelation**

In this section, I tried to filter and visualize only the statistically significant spatial clusters (p \< 0.05) from the Local Moran’s I analysis for drug use cases. This provides a more focused view of the provinces where significant clustering occurs, allowing us to pinpoint areas of interest.

### **8.4.1 Statistically Significant Spatial Autocorrelation of Drug Use Cases**

First, I filtered the **lisa_drug_use** data to include only those provinces where the p-value of Local Moran's I is less than 0.05, indicating significant local spatial autocorrelation. The map highlights these provinces and provides insight into the spatial patterns for drug use cases.

```{r}
lisa_sig_drug_use <- lisa_drug_use %>%
  filter(p_ii_sim < 0.05) %>%
  mutate(label = paste(ADM1_EN))

tm_shape(lisa_drug_use) +
  tm_polygons() +
  tm_borders(col = "black", alpha = 0.6) +
tm_shape(lisa_sig_drug_use) +
  tm_fill("ii", 
          palette = c("#b7dce9", "#e1ecbb", "#f5f3a6", "#f8d887", "#ec9a64", "#d21b1c"),
          title = "Local Moran's I (p < 0.05)",
          midpoint = NA,
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_borders(col = "black", alpha = 0.6) +
  tm_layout(main.title = "Statistically Significant Spatial Autocorrelation of Drug Use Cases",
            main.title.position = "center",
            main.title.size = 1.7,
            main.title.fontface = "bold",
            legend.title.size = 1.8,
            legend.text.size = 1.3,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", text.size = 1.5, size = 3, position = c("RIGHT", "TOP")) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), text.size = 1.2) +
  tm_grid(labels.size = 1, alpha = 0.2)
```

## **8.5 LISA Classification**

I tried to classify the spatial autocorrelation patterns using **Local Indicators of Spatial Association (LISA)**. This classification helps identify specific areas that show significant spatial relationships, including **high-high** (hotspots), **low-low** (coldspots), **high-low**, or **low-high** outlier patterns. These categories highlight areas where the spatial clustering of drug use cases is statistically significant.

### **8.5.1 Visualising LISA Classification for Drug Use Cases**

The LISA map classifies provinces based on their spatial autocorrelation categories. It uses different color codes to represent **high-high** (areas with high values surrounded by high values), **low-low** (areas with low values surrounded by low values), **high-low**, and **low-high** spatial relationships. This classification provides a clearer view of the spatial structure of drug use cases across provinces.

```{r}
tmap_mode("plot")

study_area_lisa_drug_use <- tm_shape(lisa_drug_use) +
  tm_polygons() +
  tm_borders(col = "black", alpha = 0.6) +
tm_shape(lisa_sig_drug_use) +
  tm_fill("mean", 
          palette = c("#b7dce9", "#ec9a64", "#e1ecbb", "#d21b1c"),
          title = "LISA class",
          midpoint = NA,
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_borders(col = "black", alpha = 0.6) +
  tm_layout(main.title = "Village-Level LISA Map of Drug Use Cases",
            main.title.position = "center",
            main.title.size = 1.7,
            main.title.fontface = "bold",
            legend.outside = TRUE,
            legend.outside.position = "right",
            legend.title.size = 1.8,
            legend.text.size = 1.3,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", text.size = 1.5, size = 2, position = c("RIGHT", "TOP")) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), text.size = 1.2) +
  tm_grid(labels.size = 1, alpha = 0.2)

study_area_lisa_drug_use
```

I also provided an interactive view of the LISA classification for better exploration of the data:

```{r}
tmap_mode("view")

tm_shape(lisa_drug_use) +
  tm_polygons(id = "label") +
  tm_borders(col = "black", alpha = 0.6) +
tm_shape(lisa_sig_drug_use) +
  tm_fill(
    "mean", 
    palette = c("#b7dce9", "#ec9a64", "#e1ecbb", "#d21b1c"),
    title = "LISA class",
    midpoint = NA,
    id = "label"
  ) +
  tm_borders(col = "black", alpha = 0.6)
```

**Key Observations from the Drug Use Cases LISA Map**

From the LISA classification map, I observed that several provinces showed a **low-low** pattern, meaning these provinces have low drug use cases and are surrounded by other provinces with similarly low values. The following provinces fall under the **low-low** category:

-   **Tak**

-   **Kamphaeng Phet**

-   **Nakhon Sawan**

-   **Uthai Thani**

-   **Chai Nat**

-   **Lop Buri**

-   **Sing Buri**

-   **Suphan Buri**

-   **Kanchanaburi**

-   **Ratchaburi**

-   **Phetchaburi**

-   **Prachuap Khiri Khan**

These areas can be considered **coldspots** for drug use cases.

[Continue to Page 3](Take-Home_Ex02_3.html)