---
title: "Hands On Exercise 7"
author: "Xu Haiyang"
date: "10 October, 2024" 
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

# **Overview Explanation**

In this hands-on exercise, I gained practical experience in delineating homogeneous regions using geographically referenced multivariate data. The analysis consists of two main techniques: hierarchical cluster analysis and spatially constrained cluster analysis.

### **Learning Outcome Explanation**

By completing this exercise, I learned how to:

-   Convert GIS polygon data into an R simple feature `data.frame` using appropriate functions from the `sf` package.

-   Transform the simple feature `data.frame` into an R `SpatialPolygonDataFrame` object using relevant functions from the `sf` package.

-   Perform cluster analysis using `hclust()` from Base R.

-   Execute spatially constrained cluster analysis with `skater()` from Base R.

-   Visualize the analysis output through the `ggplot2` and `tmap` packages.

### **Getting Started Explanation**

#### **Analytical Question**

In spatial policy and geobusiness, it is common to divide markets or planning areas into homogeneous regions using multivariate data. In this exercise, I focused on dividing Shan State in Myanmar into homogeneous regions using several Information and Communication Technology (ICT) indicators, including Radio, Television, Landline Phone, Mobile Phone, Computer, and Internet at home.

### **Installing and Loading R Packages Explanation**

In the code provided, I installed and loaded several R packages using `pacman::p_load()`, which efficiently manages package installation and loading:

```{r}
pacman::p_load(spdep, tmap, sf, ClustGeo, 
               ggpubr, cluster, factoextra, NbClust,
               heatmaply, corrplot, psych, tidyverse, GGally)
```

## **Data Import and Preparation Explanation**

#### **Importing Geospatial Data into R Environment**

I used the `st_read()` function from the `sf` package to read geospatial data for Myanmar township boundaries. I filtered the data to only include three regions of Shan State (East, North, and South) and selected columns 2 to 7. The resultant geospatial data was stored in `shan_sf`:

```{r}
shan_sf <- st_read(dsn = "data/geospatial", 
                   layer = "myanmar_township_boundaries") %>%
  filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)")) %>%
  select(c(2:7))
```

The dataset is then displayed using the following functions:

```{r}
shan_sf
```

```{r}
glimpse(shan_sf)
```

#### **Importing Aspatial Data into R Environment**

The aspatial dataset, which contains ICT-related data for Shan State, is imported using `read_csv()`:

```{r}
ict <- read_csv ("data/aspatial/Shan-ICT.csv")
```

```{r}
summary(ict)
```

#### **Deriving New Variables Using dplyr Package**

I created new variables using `mutate()`, calculating the proportion of households with each ICT service (per 1000 households) for Radio, Television, Landline Phone, Mobile Phone, Computer, and Internet:

```{r}
ict_derived <- ict %>%
  mutate(`RADIO_PR` = `Radio`/`Total households`*1000) %>%
  mutate(`TV_PR` = `Television`/`Total households`*1000) %>%
  mutate(`LLPHONE_PR` = `Land line phone`/`Total households`*1000) %>%
  mutate(`MPHONE_PR` = `Mobile phone`/`Total households`*1000) %>%
  mutate(`COMPUTER_PR` = `Computer`/`Total households`*1000) %>%
  mutate(`INTERNET_PR` = `Internet at home`/`Total households`*1000) %>%
  rename(`DT_PCODE` =`District Pcode`,`DT`=`District Name`,
         `TS_PCODE`=`Township Pcode`, `TS`=`Township Name`,
         `TT_HOUSEHOLDS`=`Total households`,
         `RADIO`=`Radio`, `TV`=`Television`, 
         `LLPHONE`=`Land line phone`, `MPHONE`=`Mobile phone`,
         `COMPUTER`=`Computer`, `INTERNET`=`Internet at home`) 
```

```{r}
summary(ict_derived)
```

## **Exploratory Data Analysis (EDA) Explanation**

#### **Statistical Graphics**

I used histograms and boxplots to visualize the distribution of the ICT indicators across the regions. For example, I generated a histogram for radio ownership per household:

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) +
  geom_boxplot(color="black", 
               fill="light blue")
```

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_boxplot(color="black", 
               fill="light blue")
```

```{r}
radio <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

tv <- ggplot(data=ict_derived, 
             aes(x= `TV_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

llphone <- ggplot(data=ict_derived, 
             aes(x= `LLPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

mphone <- ggplot(data=ict_derived, 
             aes(x= `MPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

computer <- ggplot(data=ict_derived, 
             aes(x= `COMPUTER_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

internet <- ggplot(data=ict_derived, 
             aes(x= `INTERNET_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

```{r}
ggarrange(radio, tv, llphone, mphone, computer, internet, 
          ncol = 3, 
          nrow = 2)
```

I applied a similar approach to plot the boxplots and histograms for other ICT indicators.

#### **Choropleth Map Creation**

I created a choropleth map by merging the spatial data (`shan_sf`) with the aspatial ICT data (`ict_derived`) using the `left_join()` function, and then wrote the updated `shan_sf` object to an RDS file:

```{r}
shan_sf <- left_join(shan_sf, 
                     ict_derived, by=c("TS_PCODE"="TS_PCODE"))
  
write_rds(shan_sf, "data/rds/shan_sf.rds")
```

```{r}
shan_sf <- read_rds("data/rds/shan_sf.rds")
```

```{r}
qtm(shan_sf, "RADIO_PR")
```

```{r}
TT_HOUSEHOLDS.map <- tm_shape(shan_sf) + 
  tm_fill(col = "TT_HOUSEHOLDS",
          n = 5,
          style = "jenks", 
          title = "Total households") + 
  tm_borders(alpha = 0.5) 

RADIO.map <- tm_shape(shan_sf) + 
  tm_fill(col = "RADIO",
          n = 5,
          style = "jenks",
          title = "Number Radio ") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(TT_HOUSEHOLDS.map, RADIO.map,
             asp=NA, ncol=2)
```

```{r}
tm_shape(shan_sf) +
    tm_polygons(c("TT_HOUSEHOLDS", "RADIO_PR"),
                style="jenks") +
    tm_facets(sync = TRUE, ncol = 2) +
  tm_legend(legend.position = c("right", "bottom"))+
  tm_layout(outer.margins=0, asp=0)
```

### **Correlation Matrix and Plot**

I calculated the correlation matrix for selected ICT indicators and created a mixed-type correlation plot using `corrplot.mixed()`. The correlation between the variables is represented with ellipses in the lower half and numeric values in the upper half.

```{r}
cluster_vars.cor = cor(ict_derived[,12:17])
corrplot.mixed(cluster_vars.cor,
         lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

### **Extracting and Cleaning Cluster Variables**

I extracted specific columns (township names and ICT indicators) from the `shan_sf` dataset, and then displayed the first 10 rows. This dataset contains the variables required for clustering.

```{r}
cluster_vars <- shan_sf %>%
  st_set_geometry(NULL) %>%
  select("TS.x", "RADIO_PR", "TV_PR", "LLPHONE_PR", "MPHONE_PR", "COMPUTER_PR")
head(cluster_vars,10)
```

```{r}
row.names(cluster_vars) <- cluster_vars$"TS.x"
head(cluster_vars,10)
```

### **Subsetting ICT Indicators and Saving/Loading Data**

I selected only the columns representing the ICT indicators and saved the resulting data frame to an RDS file for future use. I then reloaded the file to proceed with further analysis.

```{r}
shan_ict <- select(cluster_vars, c(2:6))
head(shan_ict, 10)

write_rds(shan_ict, "data/rds/shan_ict.rds")
```

```{r}
shan_ict <- read_rds("data/rds/shan_ict.rds")
```

### **Data Standardization**

I standardized the ICT data using Min-Max normalization (`normalize`) and Z-score standardization (`scale`). I also displayed summary statistics for both the normalized and Z-score standardized data.

```{r}
shan_ict.std <- normalize(shan_ict)
summary(shan_ict.std)
```

```{r}
shan_ict.z <- scale(shan_ict)
describe(shan_ict.z)
```

### **Visualization of Raw and Standardized Values**

I compared histograms of the raw ICT values and their Min-Max and Z-score standardized counterparts using `ggplot2` and arranged them side by side.

```{r}
r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Raw values without standardisation")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)
```

```{r}
r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Raw values without standardisation")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)
```

### **Hierarchical Clustering**

I computed the Euclidean distance matrix for the ICT data and performed hierarchical clustering using the Ward method. I then visualized the dendrogram.

```{r}
proxmat <- dist(shan_ict, method = 'euclidean')
```

```{r}
proxmat
```

```{r}
hclust_ward <- hclust(proxmat, method = 'ward.D')
```

```{r}
plot(hclust_ward, cex = 0.6)
```

### **Choosing the Best Clustering Method**

I calculated the agglomerative coefficient for different clustering methods (`average`, `single`, `complete`, and `ward`) to assess their performance.

```{r}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

ac <- function(x) {
  agnes(shan_ict, method = x)$ac
}

map_dbl(m, ac)
```

### **Gap Statistic for Optimal Clusters**

I applied the Gap statistic to determine the optimal number of clusters, followed by visualizing the result.

```{r}
set.seed(12345)
gap_stat <- clusGap(shan_ict, 
                    FUN = hcut, 
                    nstart = 25, 
                    K.max = 10, 
                    B = 50)
# Print the result
print(gap_stat, method = "firstmax")
```

```{r}
fviz_gap_stat(gap_stat)
```

### **Assigning Clusters and Visualizing**

I visualized the clusters identified from the hierarchical clustering and added rectangular boundaries around the clusters in the dendrogram.

```{r}
plot(hclust_ward, cex = 0.6)
rect.hclust(hclust_ward, 
            k = 6, 
            border = 2:5)
```

```{r}
shan_ict_mat <- data.matrix(shan_ict)
```

```{r}
heatmaply(normalize(shan_ict_mat),
          Colv=NA,
          dist_method = "euclidean",
          hclust_method = "ward.D",
          seriate = "OLO",
          colors = Blues,
          k_row = 6,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="Geographic Segmentation of Shan State by ICT indicators",
          xlab = "ICT Indicators",
          ylab = "Townships of Shan State"
          )
```

### **Final Cluster Assignment and Mapping**

I assigned clusters to each township and merged the cluster assignments with the original spatial data, followed by visualizing the cluster distribution on a map.

```{r}
groups <- as.factor(cutree(hclust_ward, k=6))
```

```{r}
shan_sf_cluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER`=`as.matrix.groups.`)
```

```{r}
qtm(shan_sf_cluster, "CLUSTER")
```
